---
title: Supplementary material for the paper - Fish intraspecies body size distributions.
author:
  - name: Freddie J. Heather
    orcid: 0000-0002-1650-2617
    corresponding: true
    email: Freddie.Heather@utas.edu.au
    roles:
      - Conception
      - Analysis
      - Visualisation
      - Writing
    affiliations:
      - Institute for Marine and Antarctic Studies, University of Tasmania
  - name: Shane A. Richards
    orcid: 0000-0002-9638-5827
    corresponding: false
    email: Shane.Richards@utas.edu.au
    roles:
      - Conception
      - Statistics
      - Writing
    affiliations:
      - School of Natural Sciences, University of Tasmania
  - name: Simon Brandl
    orcid: 
    corresponding: false
    email: 
    roles:
      - Data collection
    affiliations:
      - Marine Science Institute, University of Texas at Austin
  - name: Asta Audzijonyte
    orcid: 0000-0002-9919-9376
    corresponding: false
    email: Asta.Audzijonyte@utas.edu.au
    roles:
      - Conception
      - Analysis
      - Writing
    affiliations:
      - Institute for Marine and Antarctic Studies, University of Tasmania
keywords:
  - Fish body size
  - Temperature size rule
  - Empirical size distribution
  - Body size distribution
abstract: |
  Abstract goes here

format:
  pdf:
    toc: false
    number-sections: false
    colorlinks: true
  html:
    toc: true
    number-sections: false
    html-math-method: katex
  docx:
    toc: false
    number-sections: false

key-points:
  - keypoint 1
  - keypoint 2
csl: nature.csl
bibliography: references.bib
execute:
  echo: false
  warning: false
  error: false
---

# S1 - Calculating the coefficient of variation of a lognormal distribution

If body length, $L$, is log-normally distributed as:

$$
L \sim \mathcal{LN}(\nu, \tau)
$$

Then the variance of body length is:

$$
Var(L) = e^{\tau^2 -1}\cdot e^{2\nu + \tau^2}
$$

and the mean size is:

$$
Mean(L) = e^{\nu + \frac{\tau^2}{2}}
$$

then coefficient of variation is:

$$
\frac{\sqrt{Var(L)}}{Mean(L)} = \frac{\sqrt(e^{\tau^2 -1}\cdot e^{2\nu + \tau^2})}{e^{\nu + \frac{\tau^2}{2}}} = \sqrt(e^{\tau^2}-1)
$$

# S2 - 

Estimating the variation in the observed abundance within a size bin as a function of the estimated probability based on either the normal or lognormal distribution.


# S3 - Liklihood of distribution as a function of body size

```{r}


```

# S4 - Populations with a very high CV

```{r}

high_cov <- 
  plot_data %>% 
  filter(cov > 2) %>% 
  pull(population)

rls_obs_data %>% 
  filter(population %in% pp) %>% 
  ggplot() +
  aes(x = size_class, 
      y = p) +
  geom_point() + 
  geom_path() +
  facet_wrap(~population, scales = "free")

pp <- 
  plot_data %>% 
  filter(mu < 1.25 & data == "rls") %>% 
  pull(population) %>% 
  unique()

```
# S5 - Nonconvergent models

## Nonconvergent lognormal



